Exploring probability distributions for bivariate temporal granularities

### 2019-12-07 main.Rmd
estimation

\newtheorem{timedom}{Definition}[section]

\begin{timedom}\label{def:timedom}
A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{timedom}

@@

\begin{timedom}\label{def:linear}
A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:

  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
than all elements of $G(j)$, and  
  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
\end{timedom}

\begin{timedom}\label{def:granule}
Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
\end{timedom}

@@

\begin{timedom}\label{def:finerthan}
A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
for each index i, there exists an index j such that $G(i) \subset H(j)$.
\end{timedom}.

\begin{timedom}\label{def:groupsinto}
A linear granularity G groups into a linear granularity H, denoted
$G \trianglelefteq H,$
if for each index j there exists a (possibly infinite) subset S of the integers such
that
\begin{equation}
H(j) = \bigcup_{i \in S}G(i)
\end{equation}
\end{timedom}

@@

\begin{timedom}\label{def:periodical}
A granularity H is periodical with respect to a granularity G if
(1) $G \trianglelefteq H,$, and
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
$H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{timedom}

@@

\begin{timedom}\label{def:quasi}
A granularity H is quasi-periodical with respect to a granularity G if
(1) $$G \trianglelefteq H,$$, and
(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
integers R,P, where R is less than the minimum of the number of granules of H
between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
\end{timedom}

@@

\begin{timedom}\label{def:BG}
Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set. 
\end{timedom}

@@

\begin{timedom}\label{def:circular}
A circular granularity C(B, G) relates a linear granularity G to the bottom granularity B, if

\begin{equation} \label{eq:eq2}
C_{B, G}(z) = {L}(z\mod k(B,G)) \quad \forall z \in \mathbb{Z}^+
\end{equation}
where,\\
z denotes the index set, \\
B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity G with regular mapping, \\
{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
and k(B,G) is the number of granules of B in each granule of G.
\end{timedom}

**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity. However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity. Note that each
circular granularity can use a different set of labels. Example showing circular granularities relating two linear granularities each with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. G and H are two linear granularities denoting days and weeks respectively. $k(day, week) =  7$ and hence, $C_{G, B}(z)$ represented by C(BG, G) is given by $B(z\mod 7)$. The circular granularity C(B, G) consists of repetitive pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length. $L$ can be defined such that the labels corresponding to ${0,1,\dots,6}$ correspond to different days of the week ${Sun, Mon ,\dots, Sat}$ . Similarly, C(B, H) is given by $BG(z\mod 14)$ since the period length of days with respect to fortnight is 14. C(B, H) also repeats the pattern ${L'(0), \dots, L'(13)}$ in every period. Hence, circular granularities are represented as repetitive categorization of time in the diagram.

```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "}
knitr::include_graphics("Figs/circular-gran-diagram.png")
```

@@

\begin{timedom}\label{quasicircular}
A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if

@@

where,  
$z$ denotes the index set,\\
$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping,\\
{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
$T_w$ are the sets of indices of B describing $G'(w)$ such that 
\begin{equation}
G'(w) = \bigcup_{z \in T_w}B(z),\\
\end{equation}
$T_w = 0$ for $w \epsilon Z_{<0}$,\\
$\vert T_w \vert$ is the cardinality of set $T_w$.
\end{timedom}

@@

\begin{timedom}\label{aperiodic}
An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by

\begin{equation}\label{eq:eq}
\begin{split}
A_{BG, G''}(z) & = 1, z \in T_q\\
     & = 0, z \not\in T_q\\
\end{split}
\end{equation}
where,  
z denotes the index set (index of BG),
$T_q$ are the sets of indices of BG describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
\end{timedom}

@@

\begin{equation}
\begin{split}
MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
\end{split}
\end{equation}

@@

Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel. This might happen due to missing observations in the data or uneven locations of events in time domain. In such cases, the visualizations that use kernel density estimates should be used with caution as sample size would directly affect both the variance of the estimator and the confidence intervals. Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels.
the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$, unless $h$ is chosen as a function of $n$. However, the variance of the $f(x)$ is $(1/nh)\int K(t)^2dt$ is a function of n.

>>>>

\newtheorem{definition}{Definition}[section]

\begin{definition}\label{def:definition}
A time domain is a pair $(T; \le)$ where $T$ is a non-empty set of time instants and $\le$ is a total order on $T$.
\end{definition}

@@

\begin{definition}\label{def:linear}
A linear granularity is a mapping $G$ from the integers (the index set) to subsets of the time domain such that:

  (1) if $i < j$ and $G(i)$ and $G(j)$ are non-empty, then each element of $G(i)$ is less
than all elements of $G(j)$, and  
  (2) if $i < k < j$ and $G(i)$ and $G(j)$ are non-empty, then $G(k)$ is non-empty.  
\end{definition}

\begin{definition}\label{def:granule}
Each non-empty subset $G(i)$ is called a granule, where $i$ is one of the indexes and $G$ is a linear granularity.
\end{definition}

@@

\begin{definition}\label{def:finerthan}
A linear granularity G is finer than a  linear granularity H, denoted $G \preceq H$, if
for each index i, there exists an index j such that $G(i) \subset H(j)$.
\end{definition}.

\begin{definition}\label{def:groupsinto}
A linear granularity G groups into a linear granularity H, denoted
$G \trianglelefteq H,$
if for each index j there exists a (possibly infinite) subset S of the integers such
that
\begin{equation}
H(j) = \bigcup_{i \in S}G(i)
\end{equation}
\end{definition}

@@

\begin{definition}\label{def:periodical}
A granularity H is periodical with respect to a granularity G if
(1) $G \trianglelefteq H,$, and
(2) there exist $R,P \epsilon Z+$, where R is less than the number of granules of H, such
that for all $i \epsilon Z$, if $H(i) = \bigcup_{j \in S}G(j)$ and $H (i + R) \neq \phi$ then
$H (i + R) = \bigcup_{j \in S} G(j + P)$.
\end{definition}

@@

\begin{definition}\label{def:quasi}
A granularity H is quasi-periodical with respect to a granularity G if
(1) $$G \trianglelefteq H,$$, and
(2) there exist a set of intervals E1,...,Ez (the granularity exceptions) and positive
integers R,P, where R is less than the minimum of the number of granules of H
between any 2 exceptions, such that for all $i \epsilon Z$, if $H(i) = \bigcup_{k \in [0,k]} G(j_r)$  and $H (i + R) \neq \phi$ and $i + R < min(E)$, where E is the closest existing exception after $H (i)^2$, then $H (i + R) = \bigcup_{k \in [0,k]} G(j_r + P )$.
\end{definition}

@@

\begin{definition}\label{def:BG}
Bottom granularity - Given a granularity order relationship g-rel and a set of granularities having the same time domain, a granularity G in the set is a bottom granularity with respect to g-rel, if G g-rel H for each granularity H in the set. 
\end{definition}

@@

\begin{definition}\label{def:circular}
A circular granularity C(B, G) relates a linear granularity $G$ to the bottom granularity B, if

\begin{equation} \label{eq:eq5}
\begin{split}
C_{B, G}(z) & = {L}(z\mod P(B, G)) \quad \forall z \in \mathbb{Z}^+ \\
\end{split}
\end{equation}
where
z denotes the index set,
B denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G$ with regular mapping, 
$P \equiv P(B, G)$ is the number of granules of $B$ in each granule of $G$,
and ${L}$ is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (P-1)}$,

\end{definition}

**Discussion**:  In general, the set of labels will be a set of strings that is more descriptive than the index and used to identify a categorization of the circular granularity. However, the labels can coincide with indexes in which case integers are directly used to refer to categorizations of the circular granularity. Note that each circular granularity can use a different set of labels. Example showing circular granularity relating two linear granularities with bottom granularities are visually depicted in a series of slots in the \autoref{fig:circular-gran}. Each granule is represented by a box. The diagram also illustrates that the granules that overlap share elements from the underlying time domain. The first slot in the diagram shows the index set and the bottom granularity. The index of the tsibble considered is days. Note that each
circular granularity can use different label mappings. In \autoref{fig:day-of-week}, Day and Week are two linear granularities and $k(Day, Week) =  7$. The circular granularity C(Day, Week) representing Day-of-Week will consist of pattern {L(0), L(1), L(2), L(3), L(4), L(5), L(6)} which repeats itself after each period length. The label mapping $L$ can be defined as $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sun, Mon, \dots, Sat)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({Sunday, Monday, \dots, Saturday)}$ or $L: ({0,1,2, \dots, 6}) \longmapsto\ ({0, 1, \dots, 6)}$ depending on the context.

<!-- # ```{r circular-gran, echo=FALSE,out.width = "100%",fig.cap="Relating circular granularities and bottom linear granularity "} -->
<!-- # knitr::include_graphics("Figs/circular-gran-diagram.png") -->
<!-- # ``` -->

```{r day-of-week,echo=FALSE,out.width = "100%",fig.cap="Circular granularity day-of-week"}
knitr::include_graphics("Figs/day-of-week.png")
```

@@

\begin{definition}\label{quasicircular}
A quasi-circular granularity $Q_(B, G')$ that relates linear granularities $G'$ and bottom granularity $B$, if

@@

where,  
$z$ denotes the index set,\\
$B$ denotes a full-integer labelled bottom granularity which groups periodically into linear granularity $G'$ with irregular mapping,\\
{L} is a label mapping that defines an unique label for each index $l \in {0,1,\dots, (k-1)}$,\\
$T_w$ are the sets of indices of B describing $G'(w)$ such that 
\begin{equation}
G'(w) = \bigcup_{z \in T_w}B(z),\\
\end{equation}
$T_w = 0$ for $w \epsilon Z_{<0}$,\\
$\vert T_w \vert$ is the cardinality of set $T_w$.
\end{definition}

@@

\begin{definition}\label{aperiodic}
An aperiodic granularity A_(BG, G'') that relates linear granularities G and bottom granularity BG, with G'' being aperiodical with respect to BG, is given by

\begin{equation}\label{eq:eq}
\begin{split}
A_{BG, G''}(z) & = 1, z \in T_q\\
     & = 0, z \not\in T_q\\
\end{split}
\end{equation}
where,  
z denotes the index set (index of BG),
$T_q$ are the sets of indices of BG describing $G''(q)$ such that $G''(q) = \bigcup_{z \in T_q}BG(z)$,$T_q = 0$ for $q \epsilon Z_{<0}$.
\end{definition}

@@

\begin{equation}\label{eq:MSE}
\begin{split}
MSE(\hat{q(u)}) & = (1/4)h^4(q''(u))^2 \sigma_k^4 + (1/nh)q^2(u) \kappa \\
MSE(\hat{f(x)}) & = (1/4)h^4(f''(x))^2 \sigma_k^4 + (1/nh)f(x) \kappa
\end{split}
\end{equation}

@@

Even when there are no rarely occuring events, number of observations might vary hugely within or across each panel. This might happen due to missing observations in the data or uneven locations of events in time domain. In such cases, the visualizations that use density or quantile estimates should be used with caution as sample size would directly affect both the variance and consequently the confidence interval of the estimators. \autoref{eq:MSE} shows that the variance of both $\hat{f(x)}$ and $\hat{q(u)}$ is a function of the sample size $n$.

<!-- the bias in the estimation of $f(x)$ does not depend directly on the sample size($n$), but only on band width $h$,  -->

<!-- Measures such as Gini's coefficient may be used to see if sample sizes vary widely across panels. -->
